Methods to prioritise pop-up active transport infrastructure

### 2020-05-09 article.Rmd
Update paper

```{r parameters}
min_cycling_potential = 0
min_grouped_cycling_potential = 100
min_grouped_length = 500
city_centre_buffer_radius_small = 5000
city_centre_buffer_radius = 8000
city_centre_buffer_radius_large = 10000
city_centre = tmaptools::geocode_OSM("leeds", as.sf = TRUE)
city_centre_buffer_small = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius_small)
city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
city_centre_buffer_large = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius_large)
h_leeds = hsf[city_centre_buffer, ]
```

@@

To identify road sections on which there is a spare lane we developed a simple algorithm that takes the OSM variable [`lanes`](https://wiki.openstreetmap.org/wiki/Key:lanes)  if it is present and, if not, derives the number from the highway type and presence/absence of bus lanes.

<!-- Note this could be a function in an R packge.. -->

@@

```{r levels}
r_pct_lanes = rj %>% 
  filter(cycling_potential > min_cycling_potential) %>% 
  filter((lanespsvforward + lanesforward) > 1 |
      (lanespsvbackward + lanesbackward) > 1
  )
touching_list = st_touches(r_pct_lanes)
g = igraph::graph.adjlist(touching_list)
components = igraph::components(g)
r_pct_lanes$group = components$membership
r_pct_lanes = r_pct_lanes %>% 
  group_by(group) %>% 
  mutate(group_length = sum(length)) %>% 
  mutate(cycling_potential_mean = mean(cycling_potential, na.rm = TRUE)) %>% 
  filter(cycling_potential_mean > min_grouped_cycling_potential)
r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)
r_filter_before_grouping = rj %>% 
  filter(cycling_potential > min_cycling_potential) %>% 
  filter((lanespsvforward + lanesforward) > 1 |
           (lanespsvbackward + lanesbackward) > 1
  ) %>% 
  filter(cycling_potential > min_grouped_cycling_potential) %>% 
  filter(length > 100)
tmap_mode("plot")
m1 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_filter_before_grouping) + tm_lines() +
  tm_layout(title = "Filter then group (length > 100, cycling_potential > 100)")
m2 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_pct_lanes) + tm_lines() +
  tm_layout(title = "Group then filter (length > 500, cycling_potential > 100)")
ma = tmap_arrange(m1, m2)
ma
```

## Scenario development

>>>>

```{r parameters}
min_cycling_potential = 0
min_grouped_cycling_potential = 100
min_grouped_length = 600
city_centre_buffer_radius_small = 5000
city_centre_buffer_radius = 8000
city_centre_buffer_radius_large = 10000
city_centre = tmaptools::geocode_OSM("leeds", as.sf = TRUE)
city_centre_buffer_small = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius_small)
city_centre_buffer = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius)
city_centre_buffer_large = stplanr::geo_buffer(city_centre, dist = city_centre_buffer_radius_large)
h_leeds = hsf[city_centre_buffer, ]
```

@@

To identify road sections on which there is a spare lane we developed a simple algorithm that takes the OSM variable [`lanes`](https://wiki.openstreetmap.org/wiki/Key:lanes)  if it is present and, if not, derives the number from the highway type and presence/absence of bus lanes.
All segments defined as having a spare lane using this method is shown in Figure \@ref(fig:levels) (left).

<!-- Note this could be a function in an R packge.. -->
<!-- see https://github.com/cyipt/cyipt/blob/82248b2f99e388fac314d34ec5aa49bb90a737a3/scripts/prep_data/clean_osm.R#L349 -->

@@

```{r levels, fig.height=3, fig.cap="Illustration of the 'group then filter' method to identify long sections with spare lanes *and* high cycling potential"}
r_pct_lanes_all = rj %>% 
  filter(cycling_potential > min_cycling_potential) %>% 
  filter(lanes_f > 1 | lanes_b > 1)
# mapview::mapview(r_pct_lanes)

touching_list = st_touches(r_pct_lanes_all)
g = igraph::graph.adjlist(touching_list)
components = igraph::components(g)
r_pct_lanes_all$group = components$membership
r_pct_lanes = r_pct_lanes_all %>% 
  group_by(group) %>% 
  mutate(group_length = sum(length)) %>% 
  mutate(cycling_potential_mean = mean(cycling_potential, na.rm = TRUE)) %>% 
  filter(cycling_potential_mean > min_grouped_cycling_potential)
r_pct_lanes = r_pct_lanes %>% filter(group_length > min_grouped_length)
r_filter_before_grouping = rj %>% 
  filter(cycling_potential > min_cycling_potential) %>% 
  filter(lanes_f > 1 | lanes_b > 1) %>% 
  filter(cycling_potential > min_grouped_cycling_potential) %>% 
  filter(length > 100)
tmap_mode("plot")
m0 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_pct_lanes_all) + tm_lines() +
  tm_layout(title = "Roads on which there are spare lanes.")
m1 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_filter_before_grouping) + tm_lines() +
  tm_layout(title = "Filter then group:\n(length > 100, cycling_potential > 100)")
m2 = tm_shape(city_centre_buffer) + tm_borders(col = "grey") +
  tm_shape(r_pct_lanes) + tm_lines() +
  tm_layout(title = "Group then filter:\n(length > 500, cycling_potential > 100)")
# todo show group membership with colours
ma = tmap_arrange(m0, m1, m2, nrow = 1)
ma
```

## Visualisation

<!-- ## Scenario development -->
